---
title: "Rfars Package Description"
author: "Jeremy R. Winget"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rfars package description}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Package Description

The `Rfars` package reads and plots Fatality Analysis Reporting System (FARS) data 
from the [National Highway Traffic Safety 
Administration](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

[FARS](https://en.wikipedia.org/wiki/Fatality_Analysis_Reporting_System) is a
nationwide census, providing the American public yearly data, regarding fatal
injuries suffered in motor vehicle traffic crashes.

This package is primarily built for educational purposes. The package enables one to explore fatal traffic accidents from 2013-2015 using data from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (FARS).

## Load Packages

```{r}
library(Rfars)
```

## Included data

Data files included within the package are:

```{r}
list.files(system.file("extdata", package = "farsdata"))
```

## Package Functions
### Function `make_filename`

This function returns file name and path to the data files:

```{r}
make_filename(2015)
```

### Function `fars_read_years`

Auxiliary function used by `fars_summarize_years`:

```{r}
fars_read_years(2013)
```

### Function `fars_summarize_years`

This function summarizes yearly accidents data by month:

```{r}
plot(fars_summarize_years(2015))
fars_summarize_years(c(2015, 2014))
```

### Function `fars_map_state`

The function displays a state map plot including accidents location by year:

```{r}
state_num <- 49 # Utah
fars_map_state(state_num, 2014)
```
