---
title: "Rfars Package Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rfars Package Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Rfars)
```

# Using the package
## The Data

The data in this package come from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (FARS) data.

```{r sample2013_data}
fars_2013 <- make_filename(2013)
fars_2013 <- fars_read(fars_2013) 
dim(fars_2013)
fars_2013
```

For further details about the data, see the [NHTSA FARS Manuals & Documentation page](https://crashstats.nhtsa.dot.gov/#/DocumentTypeList/4).

## Loading FARS Data

To load all of the data for a given year, use the `make_filename()` and `fars_read()` functions, as shown in the [package description section](http://rpubs.com/jrwinget/package-description).

### About the Filename

Use the `make_filename` function to find out where data is stored on your machine or to create a filename to save/load data.

```{r make_filename}
fars_2013 <- make_filename(2013)
fars_2013
```

### Single Year

If you would like to examine fatality data for a single year, use the `fars_read_years()` function with a single year as input. The only data columns selected are `MONTH` and `year`. This returns a list of length one, and the first element in the list is the tibble listing the month and year for each fatal accident.

```{r single_year}
fars_2014 <- fars_read_years(years = 2014)
fars_2014[[1]]
```

### Multiple Years

If you would like to examine fatalities for multiple years, enter a vector of years as the argument for the `fars_read_years()` function (e.g., `fars_read_years(years = c(2013, 2015))` or `fars_read_years(2013:2015)`). This returns a list of tibbles, with each element of the list showing the month and year for each fatality.

```{r multiple_years}
fars_range <- fars_read_years(years = 2013:2015)
fars_range
```

## Summarizing FARS Data

The `fars_summarize_years()` function takes the same argument as the `fars_read_years()` and produces a summary of the counts of fatalities by month and year: 

```{r summarize_data}
fars_summary <- fars_summarize_years(2013:2015)
fars_summary
```

## Mapping Fatal Crashes

Finally, the `fars_map_state` function takes a state ID number and a year as arguments, and then maps that state's fatalities with a dot at the fatality location.

For a list of the state ID numbers, see page 26 of the [FARS Analytical User's Guide (2015)](https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812315#27). 

```{r mapping_crashes}
fars_map_state(53, 2014)
fars_map_state(36, 2014)
```
